---
title: "Machine Learning"
output: html_document
date: '2022-12-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("caTools")
#install.packages("FSelector")
library(FSelector)
library(tidyverse)
#install.packages("mltools")
#install.packages("tree")
library(mltools)
library(data.table)
library(caTools)
library(ROSE)
library(tree)
library(maptree)
#install.packages("partykit")
library(partykit)
#install.packages(e1071)
library(e1071)
install.packages("caret")
library(caret) 
# Packages for SVM, Random Forest and GBM
library(randomForest)
#install.packages("gbm")
library(gbm)
#load the ROCR package
#install.packages("pROC")
library(pROC) 
citation("pROC")
```
# drop duplicate data
```{r}
mydata <- read_csv("assignment_data.csv")
head(mydata,5)
mydata_distinct <- distinct(mydata)
nrow(mydata)-nrow(mydata_distinct)
```


# change data type

```{r}
LIMIT_GROUP <- cut(mydata$LIMIT,breaks = c(0,2000,5000,10000,50000,200000,1000000),labels = c("0","1","2","3","4","5"))
BILL1_GROUP <- cut(mydata$BILL1,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
BILL2_GROUP <- cut(mydata$BILL2,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
BILL3_GROUP <- cut(mydata$BILL3,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
BILL4_GROUP <- cut(mydata$BILL4,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
BILL5_GROUP <- cut(mydata$BILL5,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
BILL6_GROUP <- cut(mydata$BILL6,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT1_GROUP <- cut(mydata$PYAMT1,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT2_GROUP <- cut(mydata$PYAMT2,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT3_GROUP <- cut(mydata$PYAMT3,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT4_GROUP <- cut(mydata$PYAMT4,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT5_GROUP <- cut(mydata$PYAMT5,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
PYAMT6_GROUP <- cut(mydata$PYAMT6,breaks = c(-300000,0,2000,5000,10000,50000,2000000),labels = c("0","1","2","3","4","5"))
mydata <- cbind(mydata,LIMIT_GROUP,BILL1_GROUP,BILL2_GROUP,BILL3_GROUP,BILL4_GROUP,BILL5_GROUP,BILL6_GROUP,PYAMT1_GROUP,PYAMT2_GROUP,PYAMT3_GROUP,PYAMT4_GROUP,PYAMT5_GROUP,PYAMT6_GROUP)
columns_1 <- c("ID","GENDER","EDUCATION","MARRIAGE","AGE_CTG","PY1","PY2","PY3","PY4","PY5","PY6","SATISFACTION","FREQTRANSACTION","PHONE","DEPENDENT","RSTATUS","OTH_ACCOUNT","CAR","YEARSINADD","SECONDHOME","EMPLOYMENT","NEW_CSTM","CM_HIST","CLASS","CREDITCRD") 
columns_2 <- c("BILL1","BILL2","BILL3","BILL4","BILL5","BILL6","PYAMT1","PYAMT2","PYAMT3","PYAMT4","PYAMT5","PYAMT6","LIMIT","AGE")
columns_3 <- c("LIMIT_GROUP","BILL1_GROUP","BILL2_GROUP","BILL3_GROUP","BILL4_GROUP","BILL5_GROUP","BILL6_GROUP","PYAMT1_GROUP","PYAMT2_GROUP","PYAMT3_GROUP","PYAMT4_GROUP","PYAMT5_GROUP","PYAMT6_GROUP")
mydata[columns_1] <- lapply(mydata[columns_1],as.factor)
mydata[columns_2] <- lapply(mydata[columns_2],as.integer)
mydata[columns_3] <- lapply(mydata[columns_3],as.factor)
str(mydata)
summary(mydata)
```
# drop missing values
```{r}
mydata_new <- na.omit(mydata_distinct)
nrow(mydata_new)-nrow(mydata_distinct)
summary(mydata_distinct)
summary(mydata_new)
```
# the aggregate of delay payment(find the relationship between CLASS and PY)
```{r}
mydata_new %>% filter(PY1 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()
mydata_new %>% filter(PY2 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()
mydata_new %>% filter(PY3 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()
mydata_new %>% filter(PY4 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()
mydata_new %>% filter(PY5 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()
mydata_new %>% filter(PY6 %in%c("1","2","3","4","5","6","7","8","9")) %>% count()

```
# find the relationship between BILL and PYAMT
```{r}
mydata_new %>% mutate(relationbwbillandpat1=BILL1-PYAMT1) %>% filter(relationbwbillandpat1>0) %>% count()
mydata_new %>% mutate(relationbwbillandpat2=BILL2-PYAMT2) %>% filter(relationbwbillandpat2>0) %>% count()
mydata_new %>% mutate(relationbwbillandpat3=BILL3-PYAMT3) %>% filter(relationbwbillandpat3>0) %>% count()
mydata_new %>% mutate(relationbwbillandpat1=BILL1-PYAMT1) %>% filter(relationbwbillandpat1<=0) %>% count()
mydata_new %>% mutate(relationbwbillandpat2=BILL2-PYAMT2) %>% filter(relationbwbillandpat2<=0) %>% count()
mydata_new %>% mutate(relationbwbillandpat3=BILL3-PYAMT3) %>% filter(relationbwbillandpat3<=0) %>% count()
```


```{r}
ggplot(data = mydata_new)+geom_bar(mapping = aes(x=CLASS))
mydata_new %>% ggplot(mapping = aes(x=LIMIT))+geom_histogram()
mydata_new %>% ggplot(mapping = aes(x=BILL5))+geom_histogram()
```
# analysis the relationship between each numerical variable and CLASS
```{r}
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=LIMIT))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL1))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL2))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL3))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL4))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL5))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=BILL6))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT1))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT2))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT3))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT4))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT5))+geom_boxplot()
mydata_new %>% ggplot(mapping = aes(x=CLASS,y=PYAMT6))+geom_boxplot()
```
# analysis the relationship between each categorical variable and CLASS
```{r}
mydata %>% ggplot(mapping = aes(x=CLASS, y=GENDER))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=EDUCATION))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=MARRIAGE))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=AGE_CTG))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY1))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY2))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY3))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY4))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY5))+geom_count()
mydata %>% ggplot(mapping = aes(x=CLASS, y=PY6))+geom_count()
```


# drop no value variables
```{R}
mydata_new[c("ID","CM_HIST")] <- NULL
#"AGE", "CREDITCRD","PY2","PY3","PY4","PY5","PY6","PYAMT2","PYAMT3","PYAMT4","PYAMT5","PYAMT6","BILL2","BILL3","BILL4","BILL5","BILL6"
```
```{r}
#z_score <- function(x){
 # res <- (x-mean(x))/sd(x)
  #return(res)
#}
#newdata<- as.data.frame(sapply(mydata_new[,c("PYAMT1","PYAMT2","PYAMT3","PYAMT#4","PYAMT5","PYAMT6","BILL1","BILL2","BILL3","BILL4","BILL5","BILL6","LIMIT")]#,z_score))
#print(newdata)
#replace((unlist(mydata$c("PYAMT1","PYAMT2","PYAMT3","PYAMT4","PYAMT5","PYAMT6"#,"BILL1","BILL2","BILL3","BILL4","BILL5","BILL6","LIMIT"),newdata$c("PYAMT1","#PYAMT2","PYAMT3","PYAMT4","PYAMT5","PYAMT6","BILL1","BILL2","BILL3","BILL4","B#ILL5","BILL6","LIMIT"))

```

```{r}
#data<- cbind(newdata,mydata_new)
#data[c("LIMIT","PYAMT1","BILL1")] <- NULL
#table(data$CLASS)
str(mydata_new)
```

# partiton data 
```{r}
set.seed(123)
split = sample.split(mydata_new$CLASS, SplitRatio = 0.8)
training = subset(mydata_new, split == TRUE)
test = subset(mydata_new, split == FALSE)
table(training$CLASS)
```
```{r}
bothsampled <- ovun.sample(CLASS~.,data=training,method="both",p=0.4)$data
table(bothsampled$CLASS)
```
```{r}
#install.packages("FSelector")
library(mlbench)
attr_chi <- chi.squared(CLASS~., bothsampled)
sorted_chi <- attr_chi[order(-attr_chi$attr_importance), , drop = F]
filter(attr_chi,attr_importance>0)
```

```{r}
attr_weight <- information.gain(CLASS~., bothsampled)
print(attr_weight)
#z_score <- function(x){
#  res <- (x-mean(x))/sd(x)
#  return(res)
#}
#new_weight<- as.data.frame(sapply(attr_weight[,"attr_importance"],z_score))
#print(new_weight)
```


```{r}
sorted_weight <- attr_weight[order(-attr_weight$attr_importance), , drop = F]
print(sorted_weight)
barplot(unlist(sorted_weight),names.arg = rownames(sorted_weight),las="2",cex.names=0.5,space = 0.3)
filter(attr_weight,attr_importance>0)
str(bothsampled)
```
```{r}
#coef(mydata_new)
```

```{r}
#rcorr(as.matrix(mydata_new))
#mydata_new <- lapply(mydata_new,as.numeric)
#lm_py1 <- lm(CLASS~PY1+PY2+PY3+PY4+PY5+PY6+PYAMT1+PYAMT2+PYAMT3+PYAMT4+#PYAMT5+PYAMT6,data = mydata_new)
#summary(lm_py1)
#lm_py2 <- lm(CLASS~PY1*PY2*PY3*PY4*PY5*PY6*PYAMT1*PYAMT2*PYAMT3*PYAMT4*PYAMT5*PYAMT6,data = mydata_new)
#summary(lm_py2)
#anova(lm_py1,lm_py2)
```
```{r}
#lm_py2 <- lm(CLASS~PY1*PY2*PY3*PY4*PY5*PY6*PYAMT1*PYAMT2*PYAMT3*PYAMT4*PYAMT5*P#YAMT6,data = mydata_new)
```




```{r}
filtered_attributes <- cutoff.k(attr_weight,23)
print(filtered_attributes)
datamodelling <- training[filtered_attributes]
datamodelling$CLASS <- training$CLASS
```

#build a tree model
```{r}
set.seed(123)
dctree <- ctree(CLASS~.,datamodelling,control = tree.control(nrow(datamodelling), mindev = 0.01))
print(dctree)
plot(dctree)
```

# Predicting the Test set results 
```{r}
tree_predict <- predict(dctree,test,type = "class")
accuracy_tree <- length(which(tree_predict==test$CLASS))/nrow(test)
accuracy_tree
confusionMatrix(tree_predict,test$CLASS, positive = "1", mode = "prec_recall")
```

# Build an SVM model 
```{r}
SVM_model <- svm(CLASS~., bothsampled, kernel= "radial", scale = TRUE, probability=TRUE)
```


# Predicting the Test set results 
```{r}
svm_predict <- predict(SVM_model, test)
```


# Find the percentage of correct predictions
```{r} 
accuracy_svm <- length(which(svm_predict == test$CLASS))/nrow(test)
accuracy_svm
confusionMatrix(svm_predict,test$CLASS, positive = '1', mode = "prec_recall")

```

# Build a logistic regression model 
logistic regression model is a classification technique which models the probability that a target variable belongs to a particular class. In order to build our logistic regression model, we will use `glm()` function.

The basic syntax of this function can be given as follows:

    glm(formula, data, family = ...)
    
- Formula shows which features are used in modelling to predict target variable.

- Data is the dataset that will be used for model building.

- Family shows which type of model we want to develop. `glm()` function can be used to build generalised linear models (GLM). In order to use logistic function, we should set `family = "binomial"`.

```{r  message=FALSE}
LR_model <- glm(CLASS ~. , data = bothsampled, family = "binomial")

```


* Predict the class of the test data and store the result as *LR_prob* using `predict()` function with the following syntax:

        predict(logistic regression model, test data, type="response")

Since logistic regression returns the scores (or likelihood of belonging to a class), we use `type="response"` syntax to obtain these values.


```{r  message=FALSE}

# Predict the class probabilities of the test data
LR_prob <- predict(LR_model, test, type="response")

```

*LR_prob* will return the class scores (or probabilities). In order to predict the class of a test data, we use cutoff value 0.5. If the probability of a record is greater than or equal to 0.5, it will be marked as "1", otherwise it will be marked as "0". We need to save these predictions as factor variable.

```{r  message=FALSE}

# Predict the class 
LR_class <- ifelse(LR_prob >= 0.45, "1", "0")

# Save the predictions as factor variables
LR_class <- as.factor(LR_class)

# Find the percentage of correct predictions
accuracy_LR <- length(which(LR_class == test$class))/nrow(test)

accuracy_LR

```




#build a random forest model
```{r}

# Set random seed
set.seed(123)

# Build Random Forest model and assign it to RF_model
mtry_val <- seq(3, 7, 2)
nodesize_val <- seq(1, 10, 2)
sampsize_val <- floor(nrow(bothsampled)*c(0.5, 0.65, 0.8))
setOfvalues <- expand.grid(mtry = mtry_val, nodesize = nodesize_val, sampsize = sampsize_val)
err <- c()
# Create a data frame containing all combinations 

for (i in 1:nrow(setOfvalues)){
    # Since random forest model uses random numbers set the seed
    set.seed(123)
    
    # Train a Random Forest model
    model <- randomForest(CLASS~., bothsampled,
                          mtry = setOfvalues$mtry[i],
                          nodesize = setOfvalues$nodesize[i],
                          sampsize = setOfvalues$sampsize[i])
                          
    # Store the error rate for the model     
    err[i] <- model$err.rate[nrow(model$err.rate), "OOB"]
}

# Identify optimal set of hyperparmeters based on error rate
best_comb <- which.min(err)
print(setOfvalues[best_comb,])


```
```{r}
RF_model <- randomForest(CLASS~., bothsampled, mtry=7,nodesize=7,sampsize=15509)
importance(RF_model)


```

```{r}

# Predict the class of the test data
RF_predict <- predict(RF_model, test)
accuracy_RF <- length(which(RF_predict == test$class))/nrow(test)
accuracy_RF
confusionMatrix(RF_predict, test$CLASS, positive='1', mode = "prec_recall")

```


# build a GBM model
```{r}
datamodelling$CLASS <- as.numeric(datamodelling$CLASS)-1
```

```{r}
# Set random seed
set.seed(10)

# Build the GBM model
GBM_model <- gbm(CLASS~., datamodelling, distribution = "bernoulli", n.trees = 189, interaction.depth = 3, cv.folds = 5)

```

GBM model stops building decision trees when the number of trees reach to the limit defined by `n.trees`, which is 500 for this case. However, using all trees to make a prediction on the test data may deteriorate the performance of the model due to overfitting. We can use `gbm.perf` function to find the best number of trees to use for prediction.

```{r}
# Find the number of trees for the prediction
ntree_opt <- gbm.perf(GBM_model, method = "cv")
```

Now, we will use `predict` function to make predictions on the test data. GBM returns the probability that a target variable belongs to a particular class. Therefore, we use `type="response"` argument to obtain these values.

*GBM_prob* will keep the class scores (or probabilities). In order to predict the class of a test data, we use default threshold value. If the probability of a record is greater than or equal to 0.5, it will be marked as churn "1", otherwise it will be marked as stay "0". We need to save these predictions as factor variable.

```{r}
# Obtain prediction probabilities using ntree_opt
GBM_prob <-  predict(GBM_model, test, n.trees = ntree_opt, type = "response")

# Make predictions with threshold value 0.5
GBM_predict <- ifelse(GBM_prob>= 0.5, "1", "0")

# Save the predictions as a factor variable
GBM_predict <- as.factor(GBM_predict)

accuracy_GBM <- length(which(GBM_predict == test$CLASS))/nrow(test)
accuracy_GBM

confusionMatrix(GBM_predict, test$CLASS, positive='1', mode = "prec_recall")

```

* Compare above 5 models.

```{r}
# Return the total number of correct predictions for decision tree
#accuracy_tree

# Return the total number of correct predictions for SVM
accuracy_svm

# Return the total number of correct predictions for LR
accuracy_LR

# Return the total number of correct predictions for RF
accuracy_RF

# Return the total number of correct predictions for GBM
accuracy_GBM

# Confusion matrix
#cm_tree <- confusionMatrix(tree_predict,test$CLASS, positive = "1", mode = "prec_recall")
#cm_tree
cm_svm <- confusionMatrix(svm_predict,test$CLASS, positive = "1", mode = "prec_recall")
cm_svm
cm_lr <- confusionMatrix(LR_class,test$CLASS, positive = "1", mode = "prec_recall")
cm_lr
cm_rf <- confusionMatrix(RF_predict, test$CLASS, positive='1', mode = "prec_recall")
cm_rf
cm_GBM <- confusionMatrix(GBM_predict, test$CLASS, positive='1', mode = "prec_recall")
cm_GBM
```


Obtain class probabilities (likelihood of belonging to a class) for SVM and Random Forest models.
Class probabilities of GBM are stored in *GBM_prob*. Therefore, we only need to extract probabilities predicted by SVM and Random Forest. 
```{r}
# Obtain class probabilities by using predict() and adding type = "prob" for Random Forest
#RF_prob <- predict(RF_model, test, type = "prob")


# Add probability = TRUE for SVM
svm_prob <- predict(SVM_model, test, probability = TRUE)

# Use SVMpred to extract probabilities
SVM_prob <- attr(svm_predict, "probabilities")

```


```{r}
#ROC_RF <- roc(test$CLASS, RF_prob[,2])
ROC_SVM <- roc(test$CLASS, svm_prob[,2])
ROC_GBM <- roc(test$CLASS, GBM_prob)

# Extract required data from ROC_SVM
df_SVM = data.frame((1-ROC_SVM$specificities), ROC_SVM$sensitivities)

# Extract required data from ROC_RF
df_RF = data.frame((1-ROC_RF$specificities), ROC_RF$sensitivities)

# Extract required data from ROC_GBM
df_GBM = data.frame((1-ROC_GBM$specificities), ROC_GBM$sensitivities)

#plot the ROC curve for Random Forest, SVM and GBM
plot(df_SVM, col="red", type="l",     
xlab="False Positive Rate (1-Specificity)", ylab="True Positive Rate (Sensitivity)")
lines(df_RF, col="blue")                #adds ROC curve for RF
lines(df_GBM, col="green")              #adds ROC curve for GBM
grid(NULL, lwd = 1)

abline(a = 0, b = 1, col = "lightgray") #adds a diagonal line

legend("bottomright",
c("SVM", "Random Forest", "GBM"),
fill=c("red","blue", "green"))
```

